###Title: Exam 2017
###Author: Bongeka Buthelezi
###Date: 4 August 2017
###output: html

##Question 1
a) Multiple variables stored in one column i.e.: the group and the location should be separated 
Observations are placed in columns instead of rows (pain intensity in week 4 and week 8)
```{r echo=FALSE}
library(readr)
#import data
question_1 <- read_csv("https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2017/question_1.csv")
library(dplyr)
library(tidyr)
#tidy data
#baseline, week4 and week 8 all represent diferent pain intensities
 question_1_correct <- gather( question_1, key =Pain_Intensity, Value =c(`Week 4`, `Week 8`)) 
 #plot
```


##Question 2
#Fertility

Null:There is no significant relationship between religion and fertility
Alternative: Fertility is likely to be greatest with people who are Catholic
Type of data: Discrete, quantitative data 
Statistical test: Linear regression
Assumptions: 1. Variables are ratios
2.Linear relationship between variables
3.No outliers
4.Homoscedasticity
```{r echo= FALSE}
library(readr)
#import data
question_2 <- read_csv("https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2017/question_2.csv")
#  The data is given as ratio's e.g.: 9.96% of the polulation of courtelary is catholic while 90.04 is Protestant
# Add column to represent Protestant
library(dplyr)
library(tidyr)
question_2_correct <-mutate(question_2, Protestant = 100 - Catholic)
# Combining the two religions to form one religion variable
gather_1 <- gather(question_2_correct, key = Religion, Value = c(Catholic, Protestant))
#plot the data
plot(gather_1$value, gather_1$Fertility, main = 'religion vs fertility')
#statistical test
cor.test(gather_1$value, gather_1$Fertility,
         method = 'pearson')
```
t = -1.5261 ^-17 df= 92
p-value= 1
Interpretation: Accept H0 i.e.: The religion that a peron chooses does not affect the fertility of the person.


##Question 3
#Crop yield

Null: There is no difference in the crop yields obatined under different conditions
Alternative : Treament increases crop yield
Type of data: Continuos data
Statistical test: Repeated measures Anova
Assumptions: 1. The distribution of the data in the population is Gaussian
2. Equal variance across groups
3. The errors are independent
4. Data are unmatched
significant level= 0.05
```{r echo=FALSE}
library(dplyr)
library(tidyr)
library(readr)
#import data
question_3 <- read_csv("https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2017/question_3.csv")
#plot data
boxplot(question_3)
#statistical test
foo <- aov(control ~ treatment_1*treatment_2, data = question_3)
summary(foo)
#post hoc test
pairwise.t.test(question_3$treatment_1, question_3$treatment_2,
                p.adjust.method = 'holm',
                paired = FALSE)
```
p-value = 0.1024 (treatment_1)
0.0648 (treatmebt_2)
interpretation: accept the H0 i.e.: the condition of the field , does not affect crops yielded by that particular field


##Question 4
#Diets
Null: The type of diet the patients recieved had no influence on the overall disease outcome of the patients.
Alernative: The Mediterranean diet resulted in fewer disease outcomes.
Type of data: Catergorical, nominal data
Statistical test: Chi-squared
assumptions: 1.Randomn sampling
2. Independent observations
3. Large sample
Statistical significance = 0.05
```{r echo=FALSE}
library(dplyr)
library(tidyr)
library(readr)
#import data
question_4 <- read_csv("https://dl.dropboxusercontent.com/u/11805474/painblogr/biostats/exam-data-2017/question_4.csv")
#tidy data
#tabulate data
question_4_tabl <- xtabs(~ Diet + Condition, data = question_4)
#plot data
mosaicplot(question_4_tabl)
#statitistical test
#should use CHM however cannot install vcdExtra package
chisq.test(question_4_tabl)

```
X^2 = 15.433 df = 2
p-value = 0.0004454
Interpretation: Reject H0 , i.e.: The diet of the patient influences the overall disease outcomes with the Mediterranean being the better diet.